\documentclass[11pt,english]{article}

\usepackage[english]{babel}
\usepackage{enumitem}
\usepackage{fancyhdr}
\usepackage[top=2.3cm, bottom=2.3cm, left=1.6cm, right=1.6cm]{geometry}
\usepackage{graphicx}
\usepackage{lastpage}
\usepackage{standalone}

\setlength{\parindent}{0pt}        % indentation on new paragraph
\setlength{\parskip}{0pt}          % vertical spacing on new paragraph
\setlength{\lineskip}{1pt}         % vertical spacing between lines
\setlength{\columnsep}{1cm}        % spacing between columns
\setlength{\belowcaptionskip}{0pt} % spacing below captions
\setlength{\abovecaptionskip}{5pt} % spacong above captions

\pagestyle{fancy}

\graphicspath{{assets/}}

\lfoot{Patrick Spek}
\cfoot{\today}
\rfoot{\thepage/\pageref{LastPage}}

\newcommand*{\Frontpage}{\begingroup
	\hbox{%
		\hspace*{0.2\textwidth}
		\rule{1pt}{\textheight}
		\hspace*{0.05\textwidth}
		\parbox[b]{0.75\textwidth}{%
			{\noindent\Huge\bfseries Portfolio}\\[2\baselineskip] % Title
			{\large \textit{English writing}}\\[4\baselineskip] % Tagline or further description
			{\Large \textsc{\\
				Patrick Spek \\
			}}

			\vspace{0.5\textheight} % Whitespace between the title block and the publisher
		}
	}
\endgroup}

\renewcommand{\footrulewidth}{0.4pt}

% additional markup for included documents
\usepackage{hyperref}
\providecommand{\tightlist}{\setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\begin{document}
	% frontpage
	\thispagestyle{empty}
	\Frontpage{}

	% foreword
	\newpage
	\section{Foreword}
	This document is the portfolio for my English writing course at Avans. It
	is intended to help me learn about writing in a professional manner which I
	can use in my later career. At the start of our course I was asked to write
	down a number of goals I wanted to improve on during this course.

	\paragraph{}
	One of these goals is to improve my spelling. While there are dictionaries
	available to correct my mistakes, it would save me time and effort to
	simply do it right in a single try. This is something I wish to improve on
	during the course.

	\paragraph{}
	I tend to write a lengthy piece of text once I get started on it. However,
	this can be a chore to read for the other party, and even make the intended
	message harder to read. In turn, I want to improve on this by making
	concise messages that clearly get the intended message across, while still
	leaving enough details for the recipient to understand the reasoning behind
	the message.

	\paragraph{}
	Another of my learning goals for this class is to make less errors with
	regards to the word ordering. In formal English texts (and not-so-formal
	texts as well, for that matter), the order in which words appear is
	important. I often make mistakes in this area, for instance by putting the
	time somewhere in the middle of the sentence instead of at the end.

	\paragraph{}
	In the rest of this document you will find my results of the assignments I
	was given for this course. This includes three different documents. The
	first is an email with a complaint. The second assignment is an executive
	summary on a document made for an assignment in the first year of
	university. Lastly, there is a tutorial document. This is a rather tough
	piece of text intended for those with a little more experience in the IT
	field.

	\paragraph{}
	In addition to these finalized assignments, this portfolio contains their
	earlier versions. These were handed in to be rated by our instructor so
	that I can improve on it.

	% table of contents
	\newpage
	\tableofcontents

	% main body
	\newpage
	\section{The Email}
	\subsection{Old revision}
	\input{ewrite-email}

	\newpage
	\subsection{Final revision}
	\input{ewrite-email-new}

	\newpage
	\section{The executive summary}
	\subsection{Old revision}
	\input{ewrite-summary}

	\newpage
	\subsection{Final revision}
	\input{ewrite-summary-new}

	\newpage
	\section{The technical tutorial}
	\subsection{Old revision}
	\input{ewrite-tutorial}

	\newpage
	\subsection{Final revision}
	\input{ewrite-tutorial-new}

	\newpage
	\section{Reflection}
	The writing exercises were all pretty straightforward, yet had quite a few
	catches to improve on. My main issue was with the deadlines being
	irregular, which led to me not having feedback on the first two
	assignments. I did attend the classes, and thus did get general information
	on the subjects.

	\paragraph{}
	My main issue was with the executive summary, as this was required to be on
	a document from the previous year, which I did no longer have. I did
	receive an unfinished version from a student I did the assignment with, but
	this led to a very small summary.

	\subsection{Going forward}
	To improve further on my English skills, I intend to keep on writing
	documents in English and have it reviewed by native speakers. I have since
	published the tutorial exercise on my own website, and have plans to write
	more articles.

	Due to the technical matter of these articles, this will increase my
	English skills especially for work related discussions and presentations.
	This is in line with the required B2 level of my university, and is
	therefore exactly the kind of practice I need.
\end{document}

