\documentclass[11pt,english]{article}

\usepackage[english]{babel}

\begin{document}
  \textbf{Subject: Laptop issues}

  \paragraph{}
  Dear Sir or Madam,
\\ \\
  I am writing to inform you about issues we have with the laptops we recently
  ordered from your company, Lenovo.
\\ \\ 
  I am afraid that the batch of 600 IdeaPads we recently ordered and received
  are having technical issues which are impairing our current work flow. The
  laptops seem to have issues connecting to our WiFi network, even after we
  applied all firmware updates.
\\ \\
  Another issue we have ran into is battery failure. Even though ACPI reports
  the batteries to have well over one hour of battery life left, the machine
  shuts off and will only turn on when the laptop is connected to a power source
  or a fully charged battery is inserted.
\\ \\
  I have also received a number of complaints from my employees about the
  keyboard in the laptop. Sometimes, certain keystrokes are not being processed
  by the keyboard. We have tested this on the laptops using the xev tool, and we
  cannot figure out the reason to why it sometimes fails. Certain media keys,
  such as volume controls, do not work either.
\\ \\
  I request you to look into these issues and release an update which resolves
  them by the end of this week. If this is not possible, I wish to know your
  course of action in resolving this issue and a date at which you expect it will
  be resolved. Failure to do so will require me to send back the entire batch of
  laptops and demand a refund.

  \paragraph{}
  Yours faithfully,
  \\ \\
  Patrick Spek \\
  Chief Technical Officer, \\
  Scriptkitties
\end{document}
